---
title: "Midterm"
author: "Katie Miller"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load in packages
require(broom)
require(ggplot2)
require(plyr)
require(dplyr)
require(dplyr)
require(MASS)
require(car)
require(AER)
require(ggpubr)
require(AICcmodavg)
library(tidyverse)
library(here)
library(NatParksPalettes)
library(wesanderson)
library(lterdatasampler)
```

## Problem 1

a.  Biological: 
Ho = the phosphorus contents in the soils at the burned and unburned sites are equal. 
Ha = the soils in the burned and unburned sites do not have the same phosphorus content.
Statistical:
Ho = there is no significant difference in the mean phosphorus between the data sets.
Ha = there is a significant difference in the mean phosphorus between the data sets.

b.	A t-test can be used to determine if the mean phosphorus content is significantly different between the two areas.

c. To perform a successful student's t-test, we assume that the data sets are evenly distributed and have an equal variance.

d. If the data sets had an unequal variance we could perform a Welch's t-test.

e. The null hypothesis states that there is no difference in phosphorus content between the burned and unburned sites. 34 soil samples were collected and there was found to be a significant difference in phosphorus content between burned and unburned sites, therefore the null hypothesis was rejected (two-sample two-tailed t-test, t(66) = 2.5, ⍺ = 0.05, p = 0.014)

f. Ho = the mean phosphorus content in the soil is equal before and after the fire.
Ha = there is a difference in soil phosphorous content before and after the fire.

g.	I would perform a two-sample t-test to compare the mean phosphorus content of the soil before and after the fire in each survey site in order to determine if there is a significant difference in the means.

## Problem 2

a. Roads may hinder pollinator movement and limit pollination of plants across and along roads. 
In this study, a fluorescent pigment is used as an analogue for pollen to track the visitation of flowers by pollinators.
Generally, flowers across a road or bike path are less likely to be visited by pollinators than flowers along a road or bike path. The pedestrian path seemed to have little effect on the rate of pollination.
Habitat corridors could help pollinators cross roads and bike paths to reach flowers. Promoting alternative methods of transportation and reducing the number of cars on the road could also reduce pollinator deaths by collision and allow for safer road crossings.

b.

```{r}
# reading in data with correct file paths
sites <- read.csv(here("data", "roadsite.csv"))
fvisits <- read.csv(here("data", "flower visits.csv"))
roadcomp <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/roadcomp.csv", skip = 1)
veg <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/roadveg.csv",
                skip=1, colClasses = c(flowers = "factor",
                                       seminatural = "factor", 
                                       weed = "factor"))
veg$lawn[is.na(veg$lawn)] <- 0
veg$lawn <- as.factor(veg$lawn)

```



```{r}
roadcomp2 <- mutate(roadcomp, prop.pig.c =
                  (c.vert.n.high.pigment + c.vert.n.low.pigment)/c.vert.n.flowers,
                    prop.pig.c.hi = c.vert.n.high.pigment/c.vert.n.flowers, 
                    prop.pig.m = (m.fist.n.pigment.high + m.fist.n.pigment.low)/m.fist.n.flowers,
                  prop.pig.mi = m.fist.inflor.pigment/m.fist.n.inflorescences)
roadsite <- merge(roadcomp2, sites, by = "site")
road.c <- subset(roadsite, prop.pig.c != "NA")
road.m <- subset(roadsite, prop.pig.m != "NA")
```

```{r}
# Creating the graphs a b and c

# calculating mean, standard deviation, and standard error for figure 1a
c.prop.sum <- ddply(subset(road.c, number.of.lanes > 0), .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

# creating the graph for figure 1a 
f1a <- ggplot(c.prop.sum) + 
  # Plotting underlying data
  geom_point(data = subset(road.c, number.of.lanes > 0), 
             aes(x=position, 
                 y = prop.pig.c, 
                 shape = position,
                 color = position), 
             size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  # Plotting the mean point
  geom_point(aes(x = position, 
                 y = mean, 
                 color = position, 
                 # changed point size to better resemble figure 1
                 shape = position), size = 4) + 
  # Creating error bars for mean
  geom_errorbar(aes(ymin = mean-se, 
                    ymax = mean+se, 
                    x = position,
                    color=position), width = 0.1, linewidth = 1) + 
  # Theme and label aesthetics 
  theme_bw() + 
  # Change text to better resemble the figure 1
  theme(axis.text = element_text(colour = "black",size=rel(0.7)),
        # removed the "across" and "along" labels from x axis
        axis.text.x = element_blank(),
        legend.position = "none") + 
  labs(x = NULL,
       y = NULL)

# calculating mean, standard deviation, and standard error for figure 1b
c.prop.sum.bike <- ddply(subset(road.c, number.of.lanes == 0.5), .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

# creating the graph for figure 1b
f1b <- ggplot(c.prop.sum.bike) + 
  # Plotting underlying data
  geom_point(data = subset(road.c, number.of.lanes == 0.5),
             aes(x = position, y = prop.pig.c, shape = position, 
                 color = position), size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  # Plotting the mean point
  geom_point(aes(x=position, y = mean, color = position,
                 # changed point size to better resemble figure 1
                 shape = position), size = 4) + 
  # Creating error bars for mean
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, x = position, 
                    color = position), width = 0.1, linewidth = 1) + 
  # Theme and label aesthetics 
  theme_bw() + 
  # Changed text to better resemble the figure 1
  theme(axis.text = element_text(colour = "black",size=rel(0.7)),
        # removed the "across" and "along" labels from x axis
        axis.text.x = element_blank(),
        legend.position = "none") + 
  labs(x = NULL,
       y = NULL)


# calculating mean, standard deviation, and standard error for figure 1c
c.prop.sum.path <- ddply(subset(road.c, number.of.lanes == 0),
                         .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

# creating the graph for figure 1c
f1c <- ggplot(c.prop.sum.path) + 
  # Plotting underlying data
  geom_point(data = subset(road.c, number.of.lanes == 0),
             aes(x=position, y = prop.pig.c, shape = position, 
                 color = position), 
             size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  # Plotting the mean point
  geom_point(aes(x = position, y = mean,  color = position, 
                 # changed point size to better resemble figure 1
                 shape = position), size = 4) +
  # Creating error bars for mean
  geom_errorbar(aes(ymin = mean-se, 
                    ymax = mean+se, 
                    x = position, 
                    color = position), width = 0.1, linewidth = 1) + 
  # Theme and label aesthetics 
  theme_bw() + 
  # Changed text to better resemble the figure 1
  theme(axis.text = element_text(colour = "black",size=rel(0.7)), 
        axis.title = element_text(size=rel(0.7), face = "bold"), 
        plot.title = element_text(size = rel(0.5), face = "bold", 
                                  hjust = 0.5), 
        legend.position = "none") + 
  labs(x = "Position relative\nto pedestrian path", 
       y = NULL)

```

```{r}
# Creating graphs d and e

# calculating mean, standard deviation, and standard error for figure 1d
mi.prop.sum <- ddply(subset(road.m, number.of.lanes > 0.5), .(position), summarize,
                  mean = mean(prop.pig.mi),
                  sd = sd(prop.pig.mi),
                  se = sd(prop.pig.mi)/sqrt(length(position)))

# creating the graph for figure 1d
f1d <- ggplot(mi.prop.sum) + 
  # Plotting underlying data
  geom_point(data = subset(road.m, number.of.lanes > 0), 
             aes(x=position, y = prop.pig.mi, shape = position, color = position),
             size = 3, alpha = 0.4, position = position_jitter(width = 0.2)) +
  # Plotting the mean point
  geom_point(aes(x=position, y = mean, color=position, shape = position), size = 4) + 
  # Creating error bars for mean
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, 
                    x = position, color = position), 
                width = 0.1, linewidth = 1) + 
  # Theme and label aesthetics
  theme_bw() + 
  # change sizes of font, remove legend
  theme(axis.text = element_text(colour = "black",size=rel(0.7)), 
        axis.title = element_text(size=rel(0.7), face = "bold"), 
        plot.title = element_text(size = rel(2), face = "bold", hjust = 0.5), 
        legend.position = "none") + 
  # Label for the x axis, couldn't figure out how to do it with annotation labels 
  labs(x= "Position relative to road", y= NULL)

# calculating mean, standard deviation, and standard error for figure 1e
mi.prop.sum.bike <- ddply(subset(road.m, number.of.lanes == 0.5), .(position), summarize,
                  mean = mean(prop.pig.mi),
                  sd = sd(prop.pig.mi),
                  se = sd(prop.pig.mi)/sqrt(length(position)))

# creating the graph for figure 1d
f1e <- ggplot(mi.prop.sum.bike) + 
  # Plotting underlying data
  geom_point(data = subset(road.m, number.of.lanes == 0.5), 
             aes(x=position, y = prop.pig.mi, shape = position, color = position), 
             size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.4)) + 
  # Plotting the mean point
  geom_point(aes(x = position, y = mean, color = position, 
                 shape = position), size = 4) + 
  # Creating error bars for mean
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, x=position, color=position), 
                width = 0.1, linewidth = 1) + 
  # Theme and label aesthetics
  theme_bw() + 
  theme(axis.text = element_text(colour = "black",size=rel(0.7)), 
        axis.title = element_text(size=rel(0.7), face = "bold"), 
        plot.title = element_text(size = rel(2), face = "bold", hjust = 0.5), 
        legend.position = "none") + 
  labs(x= "Position relative\nto bike path", 
       y= NULL)

```

```{r}
# arranging the plots into a single figure
figure_1 <- ggarrange(f1a, f1b, f1c, f1d, f1e)
```

```{r}
# Adding in vertical y axis label
annotate_figure(figure_1, left = text_grob("Proportion of flowers with pigment", color = "black", rot = 90))

```


## Problem 3

a.

```{r}
# reading in data
coyotes <- read.csv(here("data",
                         "sev220_coyotescatsurvey_20150618.csv"),
                    na = c(".", ".'"))
```


```{r}
# cleaning data 
coyotes_summary <- coyotes %>%
  mutate(Habitat_type = case_when(Habitat_type == "G" ~ "grassland",
                                  Habitat_type == "S" ~ "shrubland"))
```

```{r}
# calculating the median by habitat type
coyotes_median <- coyotes_summary %>% 
  group_by(Habitat_type) %>% 
  summarize(median_l = median(Length, na.rm = TRUE))
```

```{r}
# creating the plot
ggplot(data = coyotes_median, aes(x = Habitat_type, 
                                  y = median_l, 
                                  color = Habitat_type, 
                                  shape = Habitat_type)) +
  # plotting the median point
  geom_point(size = 3, na.rm = TRUE) +
  # plotting the underlying data
  geom_jitter(data = coyotes_summary, aes(x = Habitat_type, y = Length),
              alpha = 0.3, na.rm = TRUE) +
  # labels and title
  labs(x = "Habitat Type", y = "Median Scat Length", title = "Median length of coyote scat, grassland vs. shrubland") +
  # changing theme and visuals
  theme_classic() +
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"))

```

b. 
```{r}
# visualizing missing data in coyotes data set
library(naniar)

gg_miss_var(coyotes)
```

c. Using the Mann-Whitney U test can help us determine if the median scat length is the same between the two habitat types. Our grassland and shubland data sets are random samples with similar distributions, so this test should be accurate comparing the medians of the two data sets.

d. Ho = There is no difference in median coyote scat length between the grassland and shrubland sites in the Chihuahuan Desert.
Ha = There is a difference in median coyote scat length between the grassland and shrubland sites in the Chihuahuan Desert.

e.

```{r}
# creating data frame for plot with new columns for habitat types
coyotes_test <- coyotes %>% 
  mutate(Grassland = case_when(
    Habitat_type == "G" ~ Length),
    Scrubland = case_when(
    Habitat_type == "S" ~ Length))
```


```{r}
# checking if distributions are similar in shape
ggplot(data = coyotes_summary, na.rm = TRUE) +
  stat_qq(aes(sample = Length), na.rm = TRUE) +
  stat_qq_line(aes(sample = Length), na.rm = TRUE, color = "red") +
  facet_wrap(~ Habitat_type)
```


```{r}
# Mann-Whitney U test 
wilcox.test(coyotes_test$Grassland, coyotes_test$Scrubland)
```

f. The null hypothesis is that there is no significant difference in median coyote scat length between the grassland and shrubland sites in the Chihuahuan Desert. Statistical analysis of the data sets using Wilcoxon rank-sum test shows that there is not a significant difference in the scat length and the null hypothesis is accepted (W = 35154, p-value = 0.9486)


## Problem 4

```{r}
# defining data set
icecover <- ntl_icecover
```

```{r}
# uglier plot
plot_1 <- ggplot(data = icecover, aes(x = year, y = ice_duration, color = lakeid)) +
  geom_jitter(na.rm = TRUE) +
  labs(title = "Plot 1", y = "ice duration", x = "years") +
  # built in theme
  theme_gray() +
  # change y axis label to size 10 and color dark grey
  theme(axis.text.y = element_text(colour = "darkgrey", 
                                   size = 10),
        # change x axis label to size 10 and color dark grey
        axis.text.x = element_text(colour = "darkgrey", 
                                   size = 10),
        # change y axis title to size 12 and color black
        axis.title.y = element_text(colour = "black", 
                                    size =  12),
        # change x axis title to size 12 and color black
        axis.title.x = element_text(colour = "black", 
                                    size =  12),
        # removing the legend  
        legend.position = "none",
        # changing the color and width of x axis tick marks
        axis.ticks.x = element_line(color = "black", 
                                    linewidth = 0.7),
        # changing the length and direction of x axis tick marks
        axis.ticks.length.x = unit(-.25, "cm"),
        # changing the color and width of x axis tick marks
        axis.ticks.y = element_line(color = "black", 
                                    linewidth = 0.7), 
        # changing the length and direction of x axis tick marks
        axis.ticks.length.y = unit(.25, "cm"),
        # changing y axis line color
        axis.line.y = element_line(color = "black"),
        # changing x axis line color
        axis.line.x = element_line(color = "black"),
        # changing the plot line color and line type
        panel.grid = element_line(color = "darkgrey", 
                                  linetype = 3),
        # changing plot background color
       plot.background = element_rect(fill = "white"))


   plot_1

```

b.
```{r}
# prettier plot
plot_2 <- ggplot(data = icecover, aes(x = year, y = ice_duration, color = lakeid)) +
  geom_jitter(na.rm = TRUE) +
  # adding a fit line to show decrease over time
  geom_smooth(formula = 'y ~ x', method = "lm", se = FALSE, linewidth = 0.7, na.rm = TRUE) +
  # theme and color
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.title.x = element_text(colour = "black", size = 9),
        axis.title.y = element_text(colour = "black", size = 9),
        plot.title = element_text(size =  10),
        plot.caption.position = "panel") +
  scale_color_manual(values = natparks.pals("LakeNakuru", 2)) +
  labs(y = "Ice duration (days)", x = "Year",
       title = "Days of Ice Cover at Lake Mendota and Lake Monona, 1851 - 2019",
       caption = "The duration of ice cover (in days) at Lakes Mendota and\nMonona steadily decreased between the years 1851 and 2019") 

plot_2
```

