---
title: "Midterm"
author: "Katie Miller"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(broom)
require(ggplot2)
require(plyr)
require(dplyr)
require(dplyr)
require(MASS)
require(car)
require(AER)
require(ggpubr)
require(AICcmodavg)
library(tidyverse)
library(here)
```

## Problem 1

a.  Ho = the mean phosphorus content in the burned and unburned sites is not significantly different. Ha = there is a significant difference between mean phosphorus content in the soil of the burned and unburned sites.

b.  t-test to determine if the mean of the two areas is significantly different.

c.  To perform a successful student's t-test, we assume that the data sets are evenly distributed and have an equal variance.

d.  if the data sets had an unequal variance we could perform a Welch's t-test.

e.  Because the p-value is less than 0.05, we can reject the null hypothesis that there is no difference in phosphorus content between the burned and unburned sites. 

f.  Ho = the mean phosphorus content of the soil before and after the fire is not significantly different. 
Ha = the mean phosphorus content of the soil is unchanged before and after the fires. 

g. I would preform a t-test to compare the mean phosphorus content of the soil before and after the fire in each survey site. 

## Problem 2

```{r}
# reading in data, correct file paths
roadcomp <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/roadcomp.csv", skip=1)
sites <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/roadsite.csv")
contig <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/contig.csv", colClasses = c(Site = "factor", type = "factor"))
veg <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/roadveg.csv", skip=1, colClasses = c(mowed = "factor", flowers = "factor", seminatural = "factor", weed = "factor"))
veg$lawn[is.na(veg$lawn)] <- 0
veg$lawn <- as.factor(veg$lawn)
fvisits <- read.csv("~/git/ENVS-193DS_midterm_miller_katie/data/flower visits.csv")
```

```{r}
roadcomp2 <- mutate(roadcomp, prop.pig.c = (c.vert.n.high.pigment+c.vert.n.low.pigment)/c.vert.n.flowers,
                    prop.pig.c.hi = c.vert.n.high.pigment/c.vert.n.flowers, 
                    prop.pig.m = (m.fist.n.pigment.high+m.fist.n.pigment.low)/m.fist.n.flowers, prop.pig.mi = m.fist.inflor.pigment/m.fist.n.inflorescences)
roadsite <- merge(roadcomp2, sites, by = "site")
roadveg <- merge(roadsite, veg, by = c("site", "position"))
road.c <- subset(roadsite, prop.pig.c != "NA")
road.m <- subset(roadsite, prop.pig.m != "NA")
```

```{r}
# t.test(prop.pig.c ~ position, paired = T, data = road.c)
# t.test(prop.pig.c ~ position, paired = T, data = subset(road.c, number.of.lanes > 0.5))
# t.test(prop.pig.c ~ position, paired = T, data = subset(road.c, number.of.lanes > 0))
# t.test(prop.pig.c ~ position, paired = T, data = subset(road.c, number.of.lanes == 0))
# t.test(prop.pig.c ~ position, paired = T, data = subset(road.c, number.of.lanes == 0.5))

c.prop.sum.road <- ddply(subset(road.c, number.of.lanes > 0.5), .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)),
                  n = length(position))
c.prop.sum <- ddply(subset(road.c, number.of.lanes > 0), .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

f1a <- ggplot(c.prop.sum) + 
  geom_point(data = subset(road.c, number.of.lanes > 0), 
             aes(x=position, 
                 y = prop.pig.c, 
                 shape = position,
                 color = position), 
             size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  geom_point(aes(x = position, 
                 y = mean, 
                 color = position, 
                 shape = position), size = 7) + 
  geom_errorbar(aes(ymin = mean-se, 
                    ymax = mean+se, 
                    x = position,
                    color=position), width = 0.1, size = 1) + 
  theme_bw() + 
  theme(axis.text = element_text(colour = "black",size=rel(1.3)),
        axis.title = element_text(size=rel(1.4), face = "bold"),
        legend.position = "none", plot.title = 
          element_text(size = rel(2), face = "bold", hjust = 0.5)) + 
  labs(x= NULL,
       y=NULL)

f.ga.cor <- ggplot(c.prop.sum) + 
  geom_point(data = subset(road.c, number.of.lanes > 0), 
             aes(x=position, 
                 y = prop.pig.c, 
                 shape = position, 
                 color = position), 
             size = 3, alpha = 0.4,
             position = position_jitter(width = 0.2)) + 
  geom_point(aes(x=position, y = mean, color=position, 
                 shape = position), size = 7) + 
  geom_errorbar(aes(ymin = mean-se, 
                    ymax = mean + se, 
                    x = position,
                    color=position), 
                width = 0.1, size = 1) + 
  theme_bw() + 
  theme(axis.text  =element_text(colour = "black",size=rel(1.9)),
        axis.title = element_text(size=rel(2), face = "bold"),
        legend.position = "none", 
        plot.title = element_text(size = rel(2), 
                                  face = "bold", hjust = 0.5)) +
  labs(x= NULL, 
       y= NULL)

c.prop.sum.path <- ddply(subset(road.c, number.of.lanes == 0),
                         .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

f1c <- ggplot(c.prop.sum.path) + 
  geom_point(data = subset(road.c, number.of.lanes == 0),
             aes(x=position, y = prop.pig.c, shape = position, 
                 color = position), 
             size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  geom_point(aes(x = position, 
                 y = mean, 
                 color = position, 
                 shape = position), size = 7) +
  geom_errorbar(aes(ymin = mean-se, 
                    ymax = mean+se, 
                    x = position, 
                    color = position), width = 0.1, size = 1) + 
  theme_bw() + 
  theme(axis.text = element_text(colour = "black",size=rel(1.3)), 
        axis.title = element_text(size=rel(1.4), face = "bold"), 
        plot.title = element_text(size = rel(2), face = "bold", 
                                  hjust = 0.5), 
        legend.position = "none") + 
  labs(x="Position relative\nto pedestrian path", 
       y=NULL)

c.prop.sum.bike <- ddply(subset(road.c, number.of.lanes == 0.5), .(position), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)))

f1b <- ggplot(c.prop.sum.bike) + 
  geom_point(data = subset(road.c, number.of.lanes == 0.5),
             aes(x=position, y = prop.pig.c, shape = position, 
                 color = position), size = 3, alpha = 0.4, 
             position = position_jitter(width = 0.2)) + 
  geom_point(aes(x=position, y = mean, color=position, 
                 shape = position), size = 7) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, x=position, 
                    color=position), width = 0.1, size = 1) + 
  theme_bw() + 
  theme(axis.text = element_text(colour = "black",size=rel(1.3)), 
        axis.title = element_text(size=rel(1.4), face = "bold"), 
        plot.title = element_text(size = rel(2), face = "bold",
                                  hjust = 0.5), 
        legend.position = "none") + 
  labs(x = NULL, 
       y = NULL)

c.prop.sum2 <- ddply(subset(road.c), .(position, number.of.lanes), summarize,
                  mean = mean(prop.pig.c),
                  sd = sd(prop.pig.c),
                  se = sd(prop.pig.c)/sqrt(length(position)),
                  n.zero = length(prop.pig.c[prop.pig.c == 0]),
                  n = length(position),
                  prop.zero = n.zero/n)

```

```{r}
m.prop.sum <- ddply(subset(road.m, number.of.lanes > 0.5), .(position), summarize,
                  mean = mean(prop.pig.m),
                  sd = sd(prop.pig.m),
                  se = sd(prop.pig.m)/sqrt(length(position)))
mi.prop.sum <- ddply(subset(road.m, number.of.lanes > 0.5), .(position), summarize,
                  mean = mean(prop.pig.mi),
                  sd = sd(prop.pig.mi),
                  se = sd(prop.pig.mi)/sqrt(length(position)))
f1d <- ggplot(mi.prop.sum) + geom_point(data = subset(road.m, number.of.lanes > 0), aes(x=position, y = prop.pig.mi, shape = position, color = position), size = 3, alpha = 0.4, position = position_jitter(width = 0.2)) + geom_point(aes(x=position, y = mean, color=position, shape = position), size = 7) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, x=position, color=position), width = 0.1, size = 1) + theme_bw() + theme(axis.text = element_text(colour = "black",size=rel(1.3)), axis.title = element_text(size=rel(1), face = "bold"), plot.title = element_text(size = rel(2), face = "bold", hjust = 0.5), legend.position = "none") + labs(x="Position relative to road", y="Proportion of flowers\nwith pigment")

f.ga.mon <- ggplot(mi.prop.sum) + geom_point(data = subset(road.m, number.of.lanes > 0), aes(x=position, y = prop.pig.mi, shape = position, color = position), size = 3, alpha = 0.4, position = position_jitter(width = 0.2)) + geom_point(aes(x=position, y = mean, color=position, shape = position), size = 7) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, x=position, color=position), width = 0.1, size = 1) + theme_bw() + theme(axis.text = element_text(colour = "black",size=rel(1.9)), axis.title = element_text(size=rel(2), face = "bold"), plot.title = element_text(size = rel(2), face = "bold", hjust = 0.5), legend.position = "none") + 
  labs(x="Position relative\nto road", 
       y="Proportion of flowers with pigment")

m.prop.sum.bike <- ddply(subset(road.m, number.of.lanes == 0.5), .(position), summarize,
                  mean = mean(prop.pig.m),
                  sd = sd(prop.pig.m),
                  se = sd(prop.pig.m)/sqrt(length(position)))
mi.prop.sum.bike <- ddply(subset(road.m, number.of.lanes == 0.5), .(position), summarize,
                  mean = mean(prop.pig.mi),
                  sd = sd(prop.pig.mi),
                  se = sd(prop.pig.mi)/sqrt(length(position)))
f1e <- ggplot(mi.prop.sum.bike) + geom_point(data = subset(road.m, number.of.lanes == 0.5), aes(x=position, y = prop.pig.mi, shape = position, color = position), size = 3, alpha = 0.4, position = position_jitter(width = 0.4)) + geom_point(aes(x=position, y = mean, color=position, shape = position), size = 7) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, x=position, color=position), width = 0.1, size = 1) + theme_bw() + theme(axis.text = element_text(colour = "black",size=rel(1.3)), axis.title = element_text(size=rel(1.4), face = "bold"), plot.title = element_text(size = rel(2), face = "bold", hjust = 0.5), legend.position = "none") + 
  labs(x="Position relative\nto bike path", 
       y= NULL)

m.prop.sum2 <- ddply(road.m, .(position, number.of.lanes), summarize,
                  mean = mean(prop.pig.mi),
                  sd = sd(prop.pig.mi),
                  se = sd(prop.pig.mi)/sqrt(length(position)),
                  n.zero = length(prop.pig.mi[prop.pig.mi == 0]),
                  n = length(position),
                  prop.zero = n.zero/n)
```

```{r}
ggarrange(f1a, f1b, f1c, f1d, f1e)
```



## Problem 3

a.

```{r}
coyotes <- read.csv(here("data",
                         "sev220_coyotescatsurvey_20150618.csv"),
                    na = c(".", ".'"))
```


```{r}
coyotes_summary <- coyotes %>%
  mutate(Habitat_type = case_when(Habitat_type == "G" ~ "grassland",
                                  Habitat_type == "S" ~ "shrubland"))
```

```{r}
coyotes_median <- coyotes_summary %>% 
  group_by(Habitat_type) %>% 
  summarize(median_l = median(Length, na.rm = TRUE))
```

```{r}
ggplot(data = coyotes_median, aes(x = Habitat_type, 
                                  y = median_l, 
                                  color = Habitat_type, 
                                  shape = Habitat_type)) +
  geom_point(size = 3) +
  geom_jitter(data = coyotes_summary, aes(x = Habitat_type, y = Length),
              alpha = 0.3) +
  labs(x = "Habitat Type", y = "Median Scat Length", title = "Median length of coyote scat, grassland vs. shrubland") +
  theme_classic() +
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"))

```

b. 
```{r}
library(naniar)

gg_miss_var(coyotes)
```

c.


## Problem 4

```{r}
library(lterdatasampler)
icecover <- ntl_icecover
```
a.
```{r}
# ugly plot
ggplot(data = icecover, aes(x = year, y = ice_duration)) +
  geom_jitter(na.rm = TRUE) +
  theme_bw()

```

b.
```{r}
# pretty plot
```

